\documentclass[11pt]{article}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage[backend=bibtex, style=authoryear]{biblatex}

\usepackage[mmddyyyy,hhmmss]{datetime}
\usepackage{xcolor}

\usepackage{hyperref}[hidelinks]
\definecolor{hyperblue}{rgb}{0,0,0.7}

\hypersetup{
	colorlinks = true, %Colours links instead of ugly boxes
	urlcolor = hyperblue, %Colour for external hyperlinks
	linkcolor = hyperblue, %Colour of internal links
    anchorcolor = hyperblue,
    citecolor = hyperblue,
    filecolor = hyperblue,
     }
\definecolor{nicholasCol}{RGB}{203,97,63}
\newcommand{\nicholas}[1]{{\color{nicholasCol} [\textbf{NS:} #1 (\today\ \currenttime)]}}

% \addbibresource{~/Dropbox/TAGreview/PhDreferences.bib}

\textwidth=470pt
\oddsidemargin=0pt
\topmargin=0pt
\headheight=0pt
\textheight=650pt
\headsep=0pt

\pdfpagewidth=\paperwidth
\pdfpageheight=\paperheight

\title{Statement of Research}


\author{Nicholas Santantonio}
\date{\today}


\newcommand{\gxe}{G$\times$E}
\newcommand{\gxg}{G$\times$G}


\begin{document}

\section*{\centering Statement of Research}
\begin{center} Nicholas Santantonio \end{center}

% \subsection*{Introduction}

% gxg cows
% cover crop to sequester carbon, 
% htp, 
% weather data
% crop modeling,
% abiotic biotic adaptation
% climate change
% biofuel
% sustainable cropping systems. 


% digital ag
% connect with engineering dept. 
% Fed capacity Hatch funds used only for breeding  

% Lead by Professors Royse Murphy and Don Viands, the Cornell forage breeding program has a rich legacy of providing forage crop varieties to farmers in the Northeast US, as well as collaboration nation-wide, a legacy that I hope to contribute to. 

% Forage research has been drastically under-funded compared to food crops, but renewed interest in sustainable farming has brought them back into the limelight. 

\noindent In the face of climate change, farms will have less access to agronomic inputs while they strive to maintain and update sustainable agricultural practices. This leaves genetics as the primary target for improvement of our food systems. Interest in sustainable agriculture has brought renewed attention to forage and industrial crops, which can increase soil nitrogen, sequester carbon and reduce weed populations. These crops can also be adapted to marginal lands, leaving more flexibility for farmers and less competition with food crops. Implementation of $21^{st}$ century breeding technologies will expedite improvement, but several challenges must be overcome due to the complexity of the genetics and the difficulty of phenotyping.% and costs associated with evaluation of quality.



\subsection*{Genomic selection for alfalfa}

Alfalfa, known as the ``Queen of forages'' for its high protein content, was valued in 2018 at \$127 million in NY (\href{https://www.nass.usda.gov/Quick_Stats/Ag_Overview/stateOverview.php?state=NEW%20YORK}{USDA, 2019}), and is key to the states \$2.5 billion dairy industry. Beneficial allele enrichment in alfalfa is arduous due to high inbreeding depression and the autotetraploid nature of the crop. As an obligate out-crosser, alfalfa must be bred on a population level, where varieties are released as synthetics to avoid inbreeding and exploit population-level heterosis. This has limited implementation of marker-based selection strategies because most methods are individual-based. %, seemingly requiring extensive genotyping of individuals within each alfalfa breeding population. 

% . large numbers of individuals must be genotyped and inter-mated to produce a single variety, avoid inbreeding in future generations. Further complicating the use of markers is which individual(s) should be genotyped to represent a given variety. 

% \nicholas{remove this sentence?} Generally, single individuals are not representative of the variety as a whole, and genotyping large numbers of individuals from each variety is costly and restrictive.

In pursuit of this research position, I obtained funding through a US Alfalfa Farmer Research Initiative (USAFRI) grant to develop a new population-level genomic selection framework (popGS). Tissue from many individuals within a variety or breeding population are bulked for DNA extraction and sequence-based genotyping. This allows the genomic prediction framework to operate on allele frequencies in the population as opposed to allele counts in an individual. PopGS will allow for prediction of additive effects for genetic gain, as well as dominance effects to exploit population level heterosis. This genotyping method will also be used to track allele frequency changes through time. Loci under natural selection for stand persistence will be identified and appropriately weighted in the popGS model. PopGS will require the development of affordable, medium density, sequence-based genotyping platform, presenting a potential collaboration with Breeding Insight. 

% Sufficient genotype-phenotype information will obtained by genotyping varieties and breeding material recently phenotyped by the forage breeding program, as well as all new entries moving forward. 
New populations will be formed using mathematical optimization to determine optimal contributions of parent varieties. Populations will then be planted in optimal proportions and randomly inter-mated to maximize beneficial population allele frequencies in the resulting seed lot, while minimizing inbreeding. Once shown to be effective, this strategy could revolutionize the way alfalfa seed is produced, with new varieties being defined and created by their \emph{parents}. Farmer seed would be produced by planting parental populations in optimal proportions, making new varieties rapidly available. The reduced cost of seed production could make alfalfa an affordable one or two year rotation crop for field rejuvenation.


% Once shown to be effective, this strategy could revolutionize the way alfalfa seed is produced. Instead of several generations of inter-mating to produce sufficient commercial seed, new varieties would be defined by their \emph{parents}. Whole fields could be planted in optimal parental proportions to produce farmer seed directly, thus making new varieties almost instantaneously available while maximizing the benefit of population-level heterosis. This will also reduce the cost of seed production, potentially making alfalfa an affordable one or two year rotation crop for field rejuvenation.

% allows popGS to work in a fluid fashion, where new varieties are constructed by mixing the best alleles together in optimal frequencies to leverage high additive and non-additive effects in the resulting populations. %This methodology may also be used for integration of CRISPR gene edits from single individuals into elite germplasm at high frequencies.

% CRISPR methodology is still in its infancy, but once individual plants can be transformed at all four alleles, crossing into elite germplasm will still be a challenge. Half-sib families will be at best duplex for the CRISPR allele, with further reduction in frequencies due to the inability to intracross or backcross. PopGS can be used to enrich CRISPR allele frequencies until phenotypic evaluation is feasible.

% A modified rhAmpSeq approach will be built to capture sequences likely to differ at more than one site such that multiple alleles can be tracked.  One approach may include building primer sets that are anchored in exons, but span introns, in order to maximize the likelihood of amplification and sequence differences. Specific motifs? THIS PARAGRAPH NEEDS WORK!

% In pursuit of this goal, I conceived of, wrote and was awarded a US Alfalfa Farmer Research Initiative (USAFRI) grant with Kelly Robbins, Don Viands and Julie Hansen as PI and co-PIs, respectively. Funds were procured to re-sequence eight alfalfa varieties currently under phenotypic evaluation in Geneva NY, as well as the nine historic germplasm sources of alfalfa (Segovia-Lerma et al. 2003, Jour. vol. no.). \nicholas{necessary?} A diallel population formed from the historic germplasm sources was previously developed and phenotyped in Las Cruces, NM (Segovia-Lerma 2004). These two populations will be used to augment theoretical results with empirical evidence, as well as help develop an affordable sequence-based genotyping platform, and a potential collaboration with Breeding Insight. The proposal is available on request. 


\subsection*{Developing a forage phenome}

Forages are typically harvested multiple times per year and are often perennial, requiring several years of evaluation. Forage quality must be measured to gauge animal nutrition and digestibility, but is expensive to phenotype. Leguminous cover crops are also highly effective for soil rejuvenation, but the genetic variability of these processes is largely unknown because looking underground is notoriously difficult. This makes for a heavy phenotypic burden on the breeding program that stands to benefit from high-throughput phenotyping platforms, with opportunities for collaboration with the Cornell Initiative for Digital Agriculture.

To mitigate the phenotypic burden of multiple years, cuts and the measuring of quality and root traits, multi-spectral imaging data will be collected regularly throughout each growth cycle to produce (HTP). Genetic correlations of low-throughput phenotypes (LTPs) and HTPs will be estimated and used to build genomic prediction models to predict unobserved LTPs. These HTPs will also be used to develop population specific growth curves to model development above and below ground under differing biotic and abiotic stressors. To pursue this goal, I initiated a pilot study as part of the USAFRI grant and in collaboration with the forage breeding program to evaluate the utility of multi-spectral imaging for predicting growth and biomass production.

Moving forward I intend to expand the size and number of field trials, using HTPs to limit the phenotypic burden. Instead of waiting for multi-year trial completion, rapid-cycle selection will be performed on a yearly basis using popGS and all available HTPs and LTPs from historical and in-field material. I also intend to help establish a public forage database using BreeDBase to facilitate collaboration and data sharing between forage programs.   % Models will be updated annually with HTP and LTP information from plots in the field. 
% Root phenotypes will be key to modeling carbon and nitrogen cycling, drought response, and nodulation. Tap- and fine-root architecture and turnover phenotypes can be collected through time by sampling 1 meter cores within plots, but strategies to economically select for these traits are desperately needed. 

%Tap root morphology will be measured using cores within the drill strip row, while fine root structure will be evaluated from cores sampled between drill strips within each plot. Root phenotypes will be taken twice per year, after the first and last harvests of the season, to determine root growth cycles. 

% Alfalfa cuts as lactation curves on multiple pregnancies


% These results will be used as supporting evidence for a future grant proposal to the USDA (NIFA?) to evaluate incorporation of HTP into a rapid-cycle forage breeding program. The efficacy of an HTP phenotyping strategy will be assessed by prediction of LTPs using genome wide information and HTPs. 



% selection will be assessed using several selection schemes, where crosses will be made based on genotype predictions alone, genotype plus HTP, and genotype plus HTP and LTP, along with selection on LTPs alone as a control.   
% (i.e. popGS) of cycle 0 at year 1 (C0Y1), genotype predictions including HTP at year 1, and phenotypic information from year 1 alone. After 1 round of selection, the resulting populations will be planted in the field for evaluation (C1.1). Additional phenotypes from cycle 0, year 2 (C0Y2) and cycle 1, year 1 (C1Y1) will be used to update prediction models for an additional round of selection from both C0 and C1 to produce C1.2 and C2.1 populations, respectively. Three year evaluation of these three generations will be contrasted to determine the merit of HTP assisted popGS. 

\subsection*{Synergistic systems breeding}

Forage crops are only one part of an agronomic, biological system which includes soil microorganisms, the animals that feed on the forages and the microbiome of those animals. Other than host-pathogen interactions, little attention has been paid to genomic interactions between these organisms (i.e. \gxg), despite an overall notion that they are important. These interactions can be thought of as a special case of the \gxe\ problem, where the genetic covariance of the ``environment'' (e.g. soil microorganisms) can be determined by genotyping that ``environment''. %Two interacting systems will be targeted for research into \gxg: the interaction between forage and soil microbes, and the interaction between animal and feed. 

% Interactions can then be modeled using standard \gxe\ machinery, where the genetic covariance of the biological combinations is a Kronecker product of their genetic relationships. 

Legumes can form symbiotic relationships with nitrogen fixing bacteria, \emph{Rhizobium}. Unfortunately, the signaling and infection process for nodulation is typically reduced or absent under moderate to high soil nitrogen levels. Genetic increases in nodulation could allow for the use of less chemical fertilizer, reducing the environmental impact of nitrogen runoff. I aspire to build \gxg\ prediction models for simultaneous selection of host variety and symbiont to increase the rate of colonization and nitrogen fixation, even when some nitrogen is available in the soil.

% There has been considerable research into nutrition of different feeds and how they interact with the microbiome of the rumen in dairy cows. While measures of forage quality are considered, little is known about the effect of different varieties on the animal and its microbiome. 

I also seek to establish a collaboration with the animal nutritionists, dairy farmers, and animal genetics companies to investigate the potential for synergistic forage and animal breeding. Instead of breeding animals independently of their feed, we can start to breed specialized animals to specialized feeds. Bringing animal genetics, nutrition, and farming communities together with forage breeders would set precedent for future long-term integrated breeding operations.

% Whole genome expression of host and symbiont as an intermediate interaction phenotype may also aid in understanding the genetic variation in signaling between the two species through time, under contrasting levels of available nitrogen. This may help us to understand how to select for changes in gene expression 

% targeted selection of host-symbiont combinations for testing that are likely to be beneficial, and 
% Collection, mutagenesis, and transformation of \emph{Rhizobium} may enable the genetic variation necessary 

% On the other side of the plant, 



\subsection*{Research philosophy}

In the era of big data, a shift away from small designed experiments to large observational studies at the breeding program scale is inevitable. When genotyped, stored, and made publicly available with FAIR data principals, the vast amount of data generated in a breeding program becomes a treasure trove for asking questions and informing breeding decisions. Genotyping at this scale is feasible given the drastic reduction in cost, and can be further offset by clever experimental design that trades replication at an individual level for replication at the genetic level.
% A traditional breeding program generates a vast amount of phenotypic data that is used to make yearly breeding decisions, and subsequently discarded
% This does not mean that we should cease the design and execution of experiments to address specific hypotheses, but we cannot ignore the valuable resource of observational data being collected, typically at great expense. 

I believe in a collaborative model, where breeding programs do not operate in isolation. They share germplasm, resources, expertise, and most importantly, ideas. I intend to contribute to the collaborative effort at Cornell and across the globe to build the foundational capabilities needed to deploy the latest technology for variety development. As climate change progresses, heat, drought, intense storms and hard frosts will be the new norm, and we must work together to help defend our food security through accelerated genetic improvement. 

% Sustainable here. 
% \section{other ideas}

% There is also interest in short term alfalfa crops, which can provide a large percentage of the field health gain of a full alfalfa cycle in only one or two years. This well have to be coupled with selection for high seed production under low planting density, a known a bottle neck in an alfalfa agronomic system. %There is also interest in short term alfalfa crops, which can provide a large percentage of the field health gain of a full alfalfa cycle in only one or two years. 

% \subsection*{Personnel}

% I would like to hire one post doctoral associate and one graduate student for the first year, with the potential for additional graduate students and post docs for the following year depending on grant funding. The post doc will work with me to write the popGS theory and simulation paper. The graduate student will be responsible for development of the population level genotyping platform using seed from the previously phenotyped population we select for the POC, with guidance from myself and the post doc. This will lead to two companion manuscripts, with the first on the genotyping platform, and the second on the results from the popGS POC study. 

% The forage breeding program at Cornell consists of a strong team, whose knowledge and experience will be invaluable to implementing a successful GS breeding program. Additional personnel will be brought on as necessary, including a technician to operate and manage all HTP equipment and activities.

% \subsection*{A forage community}

% One of the challenges is the lack of large phenotypic records for alfalfa in any alfalfa breeding program.

% To encourage collaboration, data sharing and a global initiative, I would like to work toward an initiative to create a North American forage database using an available platform such as T3, or BassavaBase. All tools built by our lab will be made publicly available through the (BRAPI/ galaxy etc?) to allow forage breeding programs throughout the world to make more informed mating decisions. Other programs that produce useful tools for forage breeding will also be encouraged to link them to the $\alpha\alpha$Base through BRAPI.

% % Th to encourage public (and perhaps private?) breeding programs to contribute and use data to . We want to implement a genomic prediction tool for alfalfa breeders that will also design idealized mating schemes .

% I intend to continue to foster collaborations built by Dr. Don Viands and his forage breeding research team, while hoping to expand them with new research groups, such as that of Animal Nutrition (Cornell), Ian Ray (NMSU), Charlie Brummer (UC Davis), Maria Monteros (Noble Foundation), and Debby Samac (USDA ARS at UMN, and the new forage breeder they are currently looking to hire). The forage breeding program at Cornell consists of a strong team, whose knowledge and experience will be invaluable to implementing a successful GS breeding program. Additional personnel will be brought on as necessary, including a technician to operate and manage all HTP equipment and activities.



% If proven useful, I intend to establish collaborations to implement the strategies outlined here in other forage programs in the US and abroad.    

% Finally, I recognize that the forage breeding project currently focuses on many forages other than alfalfa. It is my intent to keep those projects, while incorporating new technologies for genetic gain in these crop. My focus on alfalfa is driven by the need for reliable GS methods in a polyploid crop that must be bred at the population level. I am open to new species as well, especially if a graduate student has a desire to investigate such species.  



% The third grant proposal will be to unify breeding efforts. Genotype by genotype interactions. 

% Field establishment 


% Other long term goals include:

% Collaborate with the Agro-ecology biochemistry.  CRISPR will be used to change both the Rhizobium as well as the plant. 

% Collaboration with the animal nutrition department. sciences


% A technician will be recruited to operate the HTP platform. 

% Implementation of marker-based selection has also lagged because large numbers of individuals must be genotyped to avoid inbreeding depression in future generations. 






\end{document}
